### 2020-10-04 article.tex
draft

\documentclass[conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{silence}
  \WarningFilter*{latex}{Font shape `OT1/zi4/m/it' undefined}
\usepackage[table]{xcolor} % to highlight rows in tables
\usepackage[underline=true,rounded corners=false]{pgf-umlsd} % for UML sequence diagrams
\usepackage{inconsolata} % for nicer TT font family
\usepackage[square]{syntax} % for BNF grammar
  \setlength{\grammarparsep}{1pt}
  \setlength{\grammarindent}{2em}
\usepackage{minted} % code snippets
  \newminted{text}{linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small}
\usepackage{array} % complex structure of TABULAR-s
\usepackage{paralist} % for inline itemized lists env{inparaenum}
\usepackage{natbib} % for customization of bibliography headers
\usepackage{adjustbox}
  \newcommand{\code}[1]{\adjustbox{padding=3pt 1pt 3pt 1pt, cframe=gray!50, bgcolor=gray!10}{\texttt{#1}}}
\usepackage{tikz}
  \usetikzlibrary{shapes,arrows,shadows}
  \usetikzlibrary{decorations.pathmorphing}
  \usetikzlibrary{positioning,fit}
  \usetikzlibrary{calc}
\usepackage[hyperfootnotes=true,unicode]{hyperref} % hyper links
  \hypersetup{colorlinks=false}

\newcommand{\nospell}[1]{#1}

\begin{document}
\tolerance=1500
\raggedbottom

\title{Combining Object-Oriented Paradigm and Controlled Natural Language for Requirements Specification}
\author{Yegor Bugayenko\\yegor256@gmail.com}
\maketitle

\begin{abstract} Natural language is the dominant form of writing software requirements.
Its essential ambiguity causes inconsistency of requirements,
which leads to scope creep. On the other hand, formal requirements
specification notations such as Z, \nospell{Petri Nets}, SysML, and others are
difficult to understand by non-technical project stakeholders. They often
become a barrier between developers and requirements providers. The article
presents a controlled natural language that looks like English but
is a strongly typed object-oriented language compiled to
UML/XMI. Thus, it is easily understood, at the same time, by non-technical
people, programmers, and computers. Moreover, it is formally verifiable and
testable. It was designed, developed, and tested in three commercial
software projects in order to validate the assumption that object-oriented
design can be applied to requirements engineering at the level of
specifications writing. The article outlines key features of the language and summarizes
the experience obtained during its practical application. \end{abstract}

\section{Introduction}
\label{sec:intro}

Software Requirements Specification (SRS) should be correct, unambiguous,
complete, consistent, ranked for importance and/or stability, verifiable,
modifiable, and traceable~\cite{ieee830}. Many times, it was proven that an SRS
document that lacks some of these properties has a high likelihood of
becoming a root cause of scope and schedule problems in a
project~\cite{wiegers03,ears09,chen09}.

Natural Language (NL) is the main presentation means in industrial
requirements documents~\cite{kof10,sinha10}, mostly due to its flexibility.
Very often, contributors to requirements specifications are not experts in
requirements engineering (RE). Impreciseness and flexibility of NL, being
its advantage in informal human communications, turns into a serious drawback
in RE. Problems that appear in NL requirement documents include
ambiguity, vagueness, complexity, duplication, wordiness, and
\nospell{untestability}~\cite{ears09}.

To overcome problems associated with NL, some experts advocate the use of
other notations for the specification of requirements, such as
Z~\cite{woodcock96}, SysML, UML~\cite{uml}, Petri Nets, and others. However,
use of any of these non-textual notations often requires complex translation
of the source requirements, which can introduce further errors. Such
translation of requirements can serve to create an extra ``language
barrier'' between developers and stakeholders. There is also a training
overhead associated with the introduction of many notations~\cite{ears09}.

A desired solution to the outlined problem would be a Controlled Natural
Language (CNL), which would look like NL, but be precise,
unambiguous, and consistent. A few CNLs were introduced and developed over
the last decades. Closest to our work are Attempt to Controlled English
(ACE)~\cite{ace06}, Simplified Technical English (STE)~\cite{ste05}, Easy
Approach to Requirements Syntax (EARS)~\cite{ears09}, Boeing's Computer
Processable Language~\cite{clark05}, and \nospell{Schwitter's Processable
ENGlish}~\cite{peng}. However, there are a few critical drawbacks that exist,
to some degree, in all of these CNLs.

First, none can be mapped to objects hierarchy or a UML diagram.
They are focused more on linguistic instead of object-oriented semantic.
Some CNLs, like ACE, are based on first-order logic and can be mapped to
Lisp or other functional languages. However, as of today, object-oriented
systems dominate, especially in enterprise applications domain.

Second, they all require sequential flow of language constructs, while in
modern agile software projects wiki pages is the most popular storage for
requirements documentation, instead of files or databases. Wiki pages, by
definition, are not sorted and may contain pieces of requirements
documentation that should be combined into a complete document in a random
order.

Third, most do not allow uncertainty, while requirements development
and requirements management processes need to be able to mark
requirements as uncertain (also known as ``TBD''). Requirements engineers
need to have an ability to add a requirement into scope at one iteration
and refine it later in a few iterations. This process, called ``progressive
elaboration'' of requirements~\cite{pmbok4}, must be supported by a CNL.

To overcome these drawbacks, a new language was introduced,
implemented, and tested. Section~\ref{sec:Requs} of
the article presents the language and gives a simple example of its use,
specifying requirements of a sample software system.
Section~\ref{sec:lexical} briefly outlines lexical rules of the language.
Section~\ref{sec:syntax} describes the syntax using extended \nospell{Backus-Naur}
form and gives a few practical examples. Section~\ref{sec:semantic} outlines
a semantic layer of the compiler that is intended to detect vast majority
of inconsistencies in a requirements document.
Section~\ref{sec:uml} gives an example of to-UML mapping implemented
by the compiler. Section~\ref{sec:ambiguity}
presents a numeric metric that measures ambiguity in a requirements
specification and could be used as a guidance for system analysts.
Section~\ref{sec:results} presents empirical results obtained from three
commercial Java web projects over the last few years.
Section~\ref{sec:conclusion} presents conclusions and observations made by
the author.

\section{Controlled Natural Language}
\label{sec:Requs}

It is an object-oriented CNL that resembles English and, because of that,
can be read and understood by compilers, system analysts, end-users, and
business people. Fig.~\ref{fig:example} presents an example of a short
requirements definition document that describes a calculator that asks its
user for two floating point numbers and returns the result of their
division.

\begin{figure}
\begin{textcode}
SuD includes: user as User.
Fraction is a "math calculator".
Fraction needs:
numerator as Float, and
denominator as Float.
UC1 where SuD divides two numbers:
1. The user creates Fraction (a fraction);
2. The fraction "calculates" Float (a quotient);
3. The user "receives results" using the quotient.
UC1/2 when "division by zero":
1. The user "fails" using "can't divide by zero".
\end{textcode}
\caption{Software Requirements Specification of an example math calculator}
\label{fig:example}
\end{figure}

The example has some uncertainty expressed with plain English text sentences
in double quotes. This is an important feature that distinguishes the designed language from
other CNLs. Every class and method may be specified either formally or
by so-called ``informal'' text in double quotes. In NL requirements
specifications, such uncertainties are usually marked with TBD (``to be
determined'') placeholders~\cite{moynihan00}. Java equivalent of them are
\code{TODO} comments. Informal texts in double quotes are also used to
make references from functional requirements to customer needs, interface
specifications, supplementary documents, etc.

\code{SuD} (System under Development) is a top level class that encapsulates
all other objects and may not be ``created'' (similar to a singleton in OOP).
\code{Float} is a built-in class, as well as \code{Text}, \code{Integer}, and
a few others. Classes have methods, which are defined as use cases with a
single main flow and a few alternative flows~\cite{cockburn01}.

In the example, use case \code{UC1} is a method of class \code{SuD} and its
name is ``divides two numbers.'' Every flow step in a use case can either
\begin{inparaenum}[\itshape a\upshape)] \item call other method, or \item
create (instantiate) an object. \end{inparaenum} In the example, step
\code{1} of \code{UC1} is creating a new object \code{fraction} of class
\code{Fraction}, while step \code{2} is calling a method ``calculates'' of
\code{fraction} in order to get back an object named \code{quotient} of
class \code{Float}.

Every clause ends with a period. Order of clauses is not important
and this is yet another important feature of the language. In the example clause
at the tenth line, defining an alternative flow may appear before the use
case itself. The compiler will understand them correctly.

Fig.~\ref{fig:java} contains a close equivalent program in Java.

\begin{figure}
\inputminted[linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small]{java}{sample.java}
\caption{Java equivalent of Requs requirements document from Fig.~\ref{fig:example}}
\label{fig:java}
\end{figure}

Requs, in its current version, doesn't have any lexical/syntax constructs
for non-functional requirements (NFRs) specification. This is a subject for
future research and implementation. The biggest expected challenge in this area is
inventing (or borrowing) a method of quantification of NFRs.

\section{Lexical Analysis}
\label{sec:lexical}

There are a few main lexical terms in Requs, which can be
defined by regular expressions:

\begin{grammar}
<class> ::= `[A-Z][A-Za-z]+'

<word> ::= `[a-z]+'

<use-case-ID> ::= `UC[0-9\\.]+'

<flow-ID> ::= `[0-9]+'

<informal> ::= `"([^"]|\\")+"'
\end{grammar}

In Fig.~\ref{fig:example}, \code{SuD}, \code{Fraction}, \code{Float}, and
\code{User} are classes. They start with capital letters and contain only
letters (both lowercase and capital). \code{user}, \code{fraction}, and
\code{quotient} are variables bound to instances/objects of classes.

Besides, there are a few reserved words, which may have special meaning in
certain places of the text, including \code{the}, \code{a}, \code{includes},
\code{requires}, etc. (A full list of them is defined below in Section~\ref{sec:syntax}.)

\section{Syntax Analysis}
\label{sec:syntax}

A program consists of clauses (similar to statements in other
languages). Every clause ends with a period, like English sentences. There
are four types of clauses: class declaration, class construction, method
declaration, and alternative flow declaration.

\begin{grammar}
<SRS> ::= <clause>+

<clause> ::= <class-declaration> \\
| <class-construction> \\
| <method-declaration> \\
| <alternative-flow-declaration>
\end{grammar}

\subsection{Class Declaration}

Class declaration is a clause that declares a class as a sub-class of
another class or as a standalone class. The central part of the clause is
an \code{is a} term:

\begin{grammar}
<class-declaration> ::= <class> <is-a> <parent-class>

<parent-class> ::= <informal> | <class>

<is-a> ::= `is' ( `a' | `an' )
\end{grammar}

The second line in Fig.~\ref{fig:example} declares class \code{Fraction}
without any super-class but with an informal description \code{"math
calculator"}.

\subsection{Class Construction}

Class construction is a clause that defines arguments of a class
constructor, if necessary. By default, a constructor doesn't need any
arguments:

\begin{grammar}
<class-construction> ::= <class> <includes> `:' <slots>

<includes> ::= `includes' | `needs'

<slots> ::= <slot> ( `,' <slot> )*

<slot> ::= <variable> ( `as' <class> )?

<variable> ::= <word> `-s'?
\end{grammar}

Fig.~\ref{fig:example} defines constructor arguments only for
\code{Fraction} class. Class \code{User} don't have any arguments. Class
\code{SuD} is a pre-defined class without arguments. Class \code{Float} is a
pre-defined class with one argument.

Class without arguments in constructor normally means that it represents an
entity outside of system scope, like \code{User}.

Cardinality of associations between objects is configured by the \code{-s}
prefix at the end of variable name. There are only
two possible relationships types supported: one-to-one and one-to-many.

\subsection{Method Declaration}

Method declaration is a clause that defines a method for a class:

\begin{grammar}
<method-declaration> ::= <use-case-ID> `where' \\
<class> <signature> `:' <flows>

<signature> ::= ( <word>+ | <informal> ) \\
<subject>? <using>?

<subject> ::= <class> <binding>? | `the' <variable>

<binding> ::= `(' `a' <variable> `)'

<using> ::= `using' <subject> ( `and' <subject> )*

<flows> ::= <flow> ( `;' <flow> )+

<flow> ::= <flow-ID> `.' \\
( `The' <variable> <signature>
| `Fail' `since' <informal> )
\end{grammar}

For traceability, all methods have unique names across the entire
SRS starting with \code{UC} (use case). In other words,
methods are use cases owned by classes.

\subsection{Alternative Flow Declaration}

Alternative declaration is a clause that defines an alternative
flow of a method:

\begin{grammar}
<alternative-flow-declaration> ::= <use-case-ID> `/' \\
<flow-ID> `when' <informal> `:' <flows>
\end{grammar}

Alternative flow explains exceptional situations in one of the
flows of a method it refers to.

\section{Semantic Analysis}
\label{sec:semantic}

In the language, as in any pure object-oriented language, everything is
an object. Objects are grouped into classes. They have only methods
and no other public properties and may be bound to variables.
It is a strongly typed language.

The main goal of semantic analysis in the language is to make sure that:
\begin{inparaenum}[\itshape a\upshape)]
\item method calls match method declarations,
\item object creating have necessary constructor arguments,
\item variable bindings precede their usage,
\item methods return objects of required classes, and
\item failures are handled by alternative flows.
\end{inparaenum}
If any of these checks fail, the entire document is rejected.

\subsection{Method calls match declarations}

Every method call in every use case should either
\begin{inparaenum}[\itshape a\upshape)]
\item be informal (the 8th, 9th, and 11th lines in the example at Fig.~\ref{fig:example}),
\item formally refer to the method declared in a class, or
\item construct an instance of a class using a built-in method \code{creates} (the 7th line in the example).
\end{inparaenum}

By such a strict linking of use case flows and class methods, a consistency
of the entire document is achieved. It is no longer possible to refer to
some functionality in a use case, which doesn't exist. A rare mistake during
the requirements development phase --- but very popular during requirements
management and maintenance --- when one part of the document is updated aside
from another part.

\subsection{Object creating have necessary constructor arguments}

Every time an object is instantiated using the \code{creates} method,
the compiler checks that all required parameters of required classes are
supplied. Again, it is a very important validation that prevents inconsistency
during requirements maintenance and refinement. In the example, class
\code{Fraction} requires two parameters to be instantiated. Any attempt to
create a fraction with just one parameter or two parameters of classes other
than \code{Float} would be rejected by the compiler.

\subsection{Variable bindings precede their usage}

A variable is declared and bound to its value by means of an article
\code{a}. The seventh line of the example declares a variable
\code{fraction}, which is later used with article \code{the} (the eighth
line). If, by mistake, variable declaration is removed from the first step
of the use case or the entire step is deleted, compilation will fail.
Consistency of the document is ensured by this constraint.

\subsection{Methods return objects of required classes}

Similar to the check of parameter classes, every returned object is verified
for compliance with the class expected. The eighth line of the example
expects method ``calculates'' of class \code{Fraction} to return an object
of class \code{Float}. The method is still informal and is not defined in
the requirements document, but when a requirements expert decides to make it
formal and specify its details in a new use case, she has to make sure it
returns an instance of class \code{Float}. Otherwise, the compiler would
complain with an error.

\subsection{Failures are handled by alternative flows}

Every method in the object-oriented way may throw an exception, using a
build-in method \code{Fail since}. The method has one parameter of class
\code{Text}, which explains the reason. In the example in
Fig.~\ref{fig:example}, method ``calculates'' of class \code{Fraction} may
throw an exception and it is handled by an alternative flow. In the language, all
exceptions are checked, which means that they should be handled explicitly,
by means of alternative flows of use cases.

\section{UML and XMI}
\label{sec:uml}
\definecolor{body}{rgb}{1,1,0.8}
\definecolor{border}{rgb}{0.6039,0,0.2}
\tikzstyle{uml} = [align=left,fill=body,draw=border,font={\small\ttfamily}]
\tikzstyle{uml-class} = [uml,text ragged,rectangle split,
rectangle split parts=3,rectangle split empty part height=0em,
rectangle split every empty part={}]
\tikzstyle{comment} = [inner sep=0mm,fill=white,font={\scriptsize\ttfamily}]

When semantic analysis is done the specification looks like an object of
class \code{SuD} that encapsulates other objects and methods. This object
hierarchy is convertible to UML diagrams~\cite{uml}. Specification in
Fig.~\ref{fig:example} would be converted to the following UML diagrams
(to save space, the list is limited to the three most interesting
diagrams):

\begin{itemize}
\item Class diagram, in Fig.~\ref{fig:uml-class}
\item Object diagram, in Fig.~\ref{fig:uml-object}
\item Sequence diagram of ``divides two numbers'' method, in Fig.~\ref{fig:uml-seq}
\end{itemize}

\begin{figure}
\begin{tikzpicture}[node distance=1em]
\node[uml-class] (sud) {SuD\nodepart{third}divides two numbers};
\node[uml-class,below right=of sud] {Float\nodepart{third}};
\node[uml-class,below left=of sud] {User\nodepart{third}creates\\receives results\\fails};
\node[uml-class,below=of sud] {Fraction\nodepart{third}calculates};
\end{tikzpicture}
\caption{UML class diagram for the sample project}
\label{fig:uml-class}
\end{figure}

\begin{figure}
\begin{tikzpicture}[node distance=2.5em]
\node[uml] (sud) {\underline{:SuD}};
\node[uml,below right=of sud] (user) {\underline{:User}};
\node[uml,below left=6em of user] (fraction) {\underline{:Fraction}};
\node[uml,below left=of fraction] (numerator) {\underline{:Float}};
\node[uml,below right=of fraction] (denumerator) {\underline{:Float}};
\node[uml,left=9em of user] (quotient) {\underline{:Float}};
\path[draw,diamond-] (sud) -- node[midway,comment] {user} (user);
\path[draw,diamond-] (fraction) -- node[midway,comment] {numerator} (numerator);
\path[draw,diamond-] (fraction) -- node[midway,comment] {denumerator} (denumerator);
\path[draw,->,dashed] (user) -- node[midway,comment] {creates} (fraction);
\path[draw,->,dashed] (fraction) -- node[midway,comment] {calculates} (quotient);
\path[draw,->,dashed] (quotient) -- node[midway,comment] {receives results} (user);
\end{tikzpicture}
\caption{UML object diagram for the sample project}
\label{fig:uml-object}
\end{figure}

\begin{figure}
\begin{sequencediagram}
\tikzstyle{inststyle}=[uml,rectangle,draw,anchor=west,minimum width=1.6cm,align=center]
\newthread{user}{user}
\newinst[1]{fraction}{fraction}
\newinst[1]{quotient}{quotient}
\begin{messcall}{user}{create}{fraction}
\begin{sdblock}{try/catch}{}
\begin{call}{fraction}{calculates}{quotient}{quotient}
\end{call}
\begin{messcall}{fraction}{receives result}{user}
\end{messcall}
\begin{callself}{fraction}{division by zero}{}
\end{callself}
\begin{messcall}{fraction}{receive}{user}
\end{messcall}
\end{sdblock}
\path [draw] (0.5,-4.7) -- (7,-4.7);
\end{messcall}
\end{sequencediagram}
\caption{UML sequence diagram for ``divides two numbers'' method of class SuD of the sample project}
\label{fig:uml-seq}
\end{figure}

Every UML diagram is a formal interoperable document in XMI~\cite{xmi} that
can be rendered in web, in \LaTeX, in PDF, or translated to Java or any
other object-oriented language. Fig.~\ref{fig:xmi} shows an example XMI
for the class diagram.

\begin{figure}
\inputminted[linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small]{xml}{sample.xmi}
\caption{XMI representation of a UML diagram; a sample part}
\label{fig:xmi}
\end{figure}

\section{Ambiguity}
\label{sec:ambiguity}

With this approach, ambiguity of requirements becomes a formally measurable metric.
It is calculated as a division of informally defined methods to the total
number of methods:

\begin{equation}
A = \frac{M_{\mbox{informal}}}{M_{\mbox{total}}},\quad 0 \le A \le 1
\end{equation}

In Fig.~\ref{fig:example}, ambiguity equals to $3/4$ since there are four
methods in total and three of them are defined informally.

A method is formally defined when its signature consists of one specially
reserved verb \code{creates} and an optional list of constructor arguments,
as in the seventh line of Fig.~\ref{fig:example}. A formally defined
method doesn't have any ambiguity since it is absolutely clear what is
expected as its input and output. All it does is instantiate a new object
and bind it to a variable. In Java that would mean calling a constructor.

The higher the ambiguity, the more system analysis is required for the
requirements document to make it unambiguously understandable by everybody
in a project. The ultimate goal of a system analyst is to break down
requirements until a target ambiguity is reached. In the commercial projects
explained in Section~\ref{sec:results}, the $A$ metric was used for the planning
of RE activities. Every iteration a group of RE specialists had a goal of
decreasing the ambiguity until it reached a given value. It was not clear at
the beginning how much work that would take, but soon the RE team
obtained that knowledge and became capable of estimating
complexity of work using $A$ metric.

For example, in one of the projects, a decrease of ambiguity from $0.86$ to
$0.8$ took five working hours of a system analyst. During this work, eight
tickets were produced for discussion, each of which took from one to three
hours of work of requirements providers and other project stakeholders.
Thus, a total estimate of 0.06 decrease of ambiguity costs approximately
twenty hours of work. This is a very rough estimate and is applicable only
to the project where it was measured, but in other projects it is possible
to calculate similar metrics and manage RE activities according to them.

It was empirically observed that an SRS document becomes acceptable for
implementation and doesn't confuse programmers when its ambiguity is less
than $0.7$. Of course, it is recommended to start implementation at earlier
stages of a project, when ambiguity is rather higher, and generate requests
for the RE team to improve SRS at certain places, where ambiguity has
its peaks.

\section{Empirical Results and Future Work}
\label{sec:results}

The language was used in three commercial projects. The numbers
empirically collected are presented in Table~\ref{table:numbers}.

\begin{table}
\begin{tabular}{>{\raggedright}p{14em}rrr}
                                            & $p_1$ & $p_2$ & $p_3$ \\
\hline
Total classes                                 & 39 & 12 & 19 \\
Total methods                                 & 19 & 7 & 24 \\
Total Java classes                                  & 80 & 75 & 295 \\
\rowcolor{lightgray!50}
Ambiguity achieved                                  & 0.80 & 0.95 & 0.68 \\
Work spent on RE, staff-months                      & 0.53 & 0.21 & 0.86 \\
Total SRS contributors                              & 4 & 7 & 2 \\
Non-empty lines of Java code                        & 9.5K & 10K & 32K \\
Project duration, months                            & 7 & 26 & 11 \\
Project budget, staff-months                        & 5.5 & 7.5 & 9.5 \\
\hline
\end{tabular}
\caption{Empirically collected data from three commercial projects}
\label{table:numbers}
\end{table}

The results collected from three software projects demonstrated that
combining object-oriented paradigm with a CNL makes it possible to
specify software requirements in a predictable and verifiable manner. The
quality of requirements documentation was higher than in previously
developed projects, according to the estimates of defects reported during
project development and release phases.

It seems reasonable to analyze the effectiveness of the created CNL on
a larger number of projects, including open source ones. In order to make
such an analysis a methodology would have to be designed, to compare
the effectiveness of different methods of requirements specification
and identifying the advantages and disadvantages of each one.

\section{Conclusions}
\label{sec:conclusion}

A few important observations were made during commercial usage of the language.
First, SRS document is much shorter than discussions around it (usually kept
in bug tracking ``tickets''). It takes days to discuss one small change in a
class and just a few minutes to apply the change. Because of that, a
traceability from SRS to discussion tickets is an important feature that
has to be supported by a documentation management software. It is crucial
to have an ability to trace back every requirement and remember
the reasons behind it.

Second, complete and detailed error reporting is an important aspect of
the compiler. Initial versions of it had a technical and simplistic
errors reporting mechanism that didn't give enough information to system
analysts and business owners. They were confused trying to edit a part of
SRS in one of wiki pages and receiving a message like ``incorrect syntax on
line 45.'' Such situations required immediate attention from
programmers. Soon it became obvious that the compiler, unlike
programmers-oriented compilers of Java or C++, has to produce user friendly
error messages (and warnings). In further versions of the language, a much richer
reporting is going to be implemented.

Third, developing a requirements specification with the language requires RE
engineers to understand key principles of ``object thinking''~\cite{west04},
like ``everything is an object'' and ``objects expose behavior, not state.''
Initial training was required in two projects.

\bibliographystyle{abbrv}
\bibliography{main}

\end{document}

### 2020-10-04 sample.java

@@ -1,0 +1,29 @@
+class User {
+  public Fraction createsFraction() {
+  }
+  public void receivesResult(Float quotient) {
+  }
+  public void fails(String message) {
+  }
+}
+class Fraction {
+  private final float numerator;
+  private final float denumerator;
+  public Fraction(float num, float denum) {
+    this.numerator = num;
+    this.denumerator = denum;
+  }
+  public Float quotient() throws DivisionByZero {
+    }
+  }
+  class SuD {
+    private final User user;
+    public dividesTwoNumbers() {
+    Fraction fraction = this.user.createsFraction();
+    try {
+      this.user.receivesResult(fraction.quotient());
+    } catch (DivisionByZero ex) {
+      this.user.fails("denominator can't be zero");
+    }
+  }
+}

### 2021-08-01 article.tex
acmart

@@
-\documentclass[conference]{IEEEtran}
-\usepackage[utf8]{inputenc}
-\usepackage{silence}
-  \WarningFilter*{latex}{Font shape `OT1/zi4/m/it' undefined}
-\usepackage[table]{xcolor} % to highlight rows in tables
+\documentclass[sigplan,10pt,nonacm=true]{acmart}
+\settopmatter{printfolios=false,printccs=false,printacmref=false}
 \usepackage[underline=true,rounded corners=false]{pgf-umlsd} % for UML sequence diagrams
-\usepackage{inconsolata} % for nicer TT font family
 \usepackage[square]{syntax} % for BNF grammar
@@
   \setlength{\grammarindent}{2em}
-\usepackage{minted} % code snippets
-  \newminted{text}{linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small}
 \usepackage{array} % complex structure of TABULAR-s
@@
 \usepackage{natbib} % for customization of bibliography headers
-\usepackage{adjustbox}
-  \newcommand{\code}[1]{\adjustbox{padding=3pt 1pt 3pt 1pt, cframe=gray!50, bgcolor=gray!10}{\texttt{#1}}}
+\usepackage{ffcode}
 \usepackage{tikz}
@@
   \usetikzlibrary{calc}
-\usepackage[hyperfootnotes=true,unicode]{hyperref} % hyper links
-  \hypersetup{colorlinks=false}
+\usepackage{href-ul} % hyper links
 
@@
 \begin{document}
-\tolerance=1500
-\raggedbottom
 
@@
 \author{Yegor Bugayenko\\yegor256@gmail.com}
-\maketitle
-
 \begin{abstract} Natural language is the dominant form of writing software requirements.
@@
 the experience obtained during its practical application. \end{abstract}
+\maketitle
 
@@
 complete, consistent, ranked for importance and/or stability, verifiable,
-modifiable, and traceable~\cite{ieee830}. Many times, it was proven that an SRS
+modifiable, and traceable~\citep{ieee830}. Many times, it was proven that an SRS
 document that lacks some of these properties has a high likelihood of
 becoming a root cause of scope and schedule problems in a
-project~\cite{wiegers03,ears09,chen09}.
+project~\citep{wiegers03,ears09,chen09}.
 
 Natural Language (NL) is the main presentation means in industrial
-requirements documents~\cite{kof10,sinha10}, mostly due to its flexibility.
+requirements documents~\citep{kof10,sinha10}, mostly due to its flexibility.
 Very often, contributors to requirements specifications are not experts in
@@
 ambiguity, vagueness, complexity, duplication, wordiness, and
-\nospell{untestability}~\cite{ears09}.
+\nospell{untestability}~\citep{ears09}.
 
@@
 other notations for the specification of requirements, such as
-Z~\cite{woodcock96}, SysML, UML~\cite{uml}, Petri Nets, and others. However,
+Z~\citep{woodcock96}, SysML, UML~\citep{uml}, Petri Nets, and others. However,
 use of any of these non-textual notations often requires complex translation
@@
 barrier'' between developers and stakeholders. There is also a training
-overhead associated with the introduction of many notations~\cite{ears09}.
+overhead associated with the introduction of many notations~\citep{ears09}.
 
@@
 the last decades. Closest to our work are Attempt to Controlled English
-(ACE)~\cite{ace06}, Simplified Technical English (STE)~\cite{ste05}, Easy
-Approach to Requirements Syntax (EARS)~\cite{ears09}, Boeing's Computer
-Processable Language~\cite{clark05}, and \nospell{Schwitter's Processable
-ENGlish}~\cite{peng}. However, there are a few critical drawbacks that exist,
+(ACE)~\citep{ace06}, Simplified Technical English (STE)~\citep{ste05}, Easy
+Approach to Requirements Syntax (EARS)~\citep{ears09}, Boeing's Computer
+Processable Language~\citep{clark05}, and \nospell{Schwitter's Processable
+ENGlish}~\citep{peng}. However, there are a few critical drawbacks that exist,
 to some degree, in all of these CNLs.
@@
 and refine it later in a few iterations. This process, called ``progressive
-elaboration'' of requirements~\cite{pmbok4}, must be supported by a CNL.
+elaboration'' of requirements~\citep{pmbok4}, must be supported by a CNL.
 
@@
 can be read and understood by compilers, system analysts, end-users, and
-business people. Fig.~\ref{fig:example} presents an example of a short
+business people. The Fig.~\ref{fig:example} presents an example of a short
 requirements definition document that describes a calculator that asks its
@@
 \begin{figure}
-\begin{textcode}
+\begin{ffcode}
 SuD includes: user as User.
@@
 1. The user creates Fraction (a fraction);
-2. The fraction "calculates" Float (a quotient);
-3. The user "receives results" using the quotient.
+2. The fraction "calculates" Float
+  (a quotient);
+3. The user "receives results" using
+  the quotient.
 UC1/2 when "division by zero":
-1. The user "fails" using "can't divide by zero".
-\end{textcode}
+1. The user "fails" using "can't
+  divide by zero".
+\end{ffcode}
 \caption{Software Requirements Specification of an example math calculator}
@@
 specifications, such uncertainties are usually marked with TBD (``to be
-determined'') placeholders~\cite{moynihan00}. Java equivalent of them are
-\code{TODO} comments. Informal texts in double quotes are also used to
+determined'') placeholders~\citep{moynihan00}. Java equivalent of them are
+\ff{TODO} comments. Informal texts in double quotes are also used to
 make references from functional requirements to customer needs, interface
@@
 
-\code{SuD} (System under Development) is a top level class that encapsulates
+\ff{SuD} (System under Development) is a top level class that encapsulates
 all other objects and may not be ``created'' (similar to a singleton in OOP).
-\code{Float} is a built-in class, as well as \code{Text}, \code{Integer}, and
+\ff{Float} is a built-in class, as well as \ff{Text}, \ff{Integer}, and
 a few others. Classes have methods, which are defined as use cases with a
-single main flow and a few alternative flows~\cite{cockburn01}.
+single main flow and a few alternative flows~\citep{cockburn01}.
 
-In the example, use case \code{UC1} is a method of class \code{SuD} and its
+In the example, use case \ff{UC1} is a method of class \ff{SuD} and its
 name is ``divides two numbers.'' Every flow step in a use case can either
@@
 create (instantiate) an object. \end{inparaenum} In the example, step
-\code{1} of \code{UC1} is creating a new object \code{fraction} of class
-\code{Fraction}, while step \code{2} is calling a method ``calculates'' of
-\code{fraction} in order to get back an object named \code{quotient} of
-class \code{Float}.
+\ff{1} of \ff{UC1} is creating a new object \ff{fraction} of class
+\ff{Fraction}, while step \ff{2} is calling a method ``calculates'' of
+\ff{fraction} in order to get back an object named \ff{quotient} of
+class \ff{Float}.
 
@@
 
-Fig.~\ref{fig:java} contains a close equivalent program in Java.
+The Fig.~\ref{fig:java} contains a close equivalent program in Java.
 
 \begin{figure}
-\inputminted[linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small]{java}{sample.java}
-\caption{Java equivalent of Requs requirements document from Fig.~\ref{fig:example}}
+\begin{ffcode}
+class User {
+  Fraction createsFraction() {
+  }
+  void receivesResult(Float quotient) {
+  }
+  void fails(String message) {
+  }
+}
+class Fraction {
+  private final float numerator;
+  private final float denumerator;
+  Fraction(float num, float denum) {
+    this.numerator = num;
+    this.denumerator = denum;
+  }
+  Float quotient() throws DivisionByZero {
+  }
+}
+class SuD {
+  private final User user;
+  void dividesTwoNumbers() {
+    Fraction fraction = this.user.createsFraction();
+    try {
+      this.user.receivesResult(
+        fraction.quotient()
+      );
+    } catch (DivisionByZero ex) {
+      this.user.fails(
+        "denominator can't be zero"
+      );
+    }
+  }
+}
+\end{ffcode}
+\caption{Java equivalent of Requs requirements document from the Fig.~\ref{fig:example}}
 \label{fig:java}
@@
 
-In Fig.~\ref{fig:example}, \code{SuD}, \code{Fraction}, \code{Float}, and
-\code{User} are classes. They start with capital letters and contain only
-letters (both lowercase and capital). \code{user}, \code{fraction}, and
-\code{quotient} are variables bound to instances/objects of classes.
+In the Fig.~\ref{fig:example}, \ff{SuD}, \ff{Fraction}, \ff{Float}, and
+\ff{User} are classes. They start with capital letters and contain only
+letters (both lowercase and capital). \ff{user}, \ff{fraction}, and
+\ff{quotient} are variables bound to instances/objects of classes.
 
 Besides, there are a few reserved words, which may have special meaning in
-certain places of the text, including \code{the}, \code{a}, \code{includes},
-\code{requires}, etc. (A full list of them is defined below in Section~\ref{sec:syntax}.)
+certain places of the text, including \ff{the}, \ff{a}, \ff{includes},
+\ff{requires}, etc. (A full list of them is defined below in Section~\ref{sec:syntax}.)
 
@@
 another class or as a standalone class. The central part of the clause is
-an \code{is a} term:
+an \ff{is a} term:
 
@@
 
-The second line in Fig.~\ref{fig:example} declares class \code{Fraction}
-without any super-class but with an informal description \code{"math
-calculator"}.
+The second line in the the Fig.~\ref{fig:example} declares class \ff{Fraction}
+without any super-class but with an informal description
+\ff{"math calculator"}.
 
@@
 
-Fig.~\ref{fig:example} defines constructor arguments only for
-\code{Fraction} class. Class \code{User} don't have any arguments. Class
-\code{SuD} is a pre-defined class without arguments. Class \code{Float} is a
+The Fig.~\ref{fig:example} defines constructor arguments only for
+\ff{Fraction} class. Class \ff{User} don't have any arguments. Class
+\ff{SuD} is a pre-defined class without arguments. Class \ff{Float} is a
 pre-defined class with one argument.
@@
 Class without arguments in constructor normally means that it represents an
-entity outside of system scope, like \code{User}.
+entity outside of system scope, like \ff{User}.
 
-Cardinality of associations between objects is configured by the \code{-s}
+Cardinality of associations between objects is configured by the \ff{-s}
 prefix at the end of variable name. There are only
@@
 For traceability, all methods have unique names across the entire
-SRS starting with \code{UC} (use case). In other words,
+SRS starting with \ff{UC} (use case). In other words,
 methods are use cases owned by classes.
@@
 \begin{inparaenum}[\itshape a\upshape)]
-\item be informal (the 8th, 9th, and 11th lines in the example at Fig.~\ref{fig:example}),
+\item be informal (the 8th, 9th, and 11th lines in the example at the Fig.~\ref{fig:example}),
 \item formally refer to the method declared in a class, or
-\item construct an instance of a class using a built-in method \code{creates} (the 7th line in the example).
+\item construct an instance of a class using a built-in method \ff{creates} (the 7th line in the example).
 \end{inparaenum}
@@
 
-Every time an object is instantiated using the \code{creates} method,
+Every time an object is instantiated using the \ff{creates} method,
 the compiler checks that all required parameters of required classes are
@@
 during requirements maintenance and refinement. In the example, class
-\code{Fraction} requires two parameters to be instantiated. Any attempt to
+\ff{Fraction} requires two parameters to be instantiated. Any attempt to
 create a fraction with just one parameter or two parameters of classes other
-than \code{Float} would be rejected by the compiler.
+than \ff{Float} would be rejected by the compiler.
 
@@
 A variable is declared and bound to its value by means of an article
-\code{a}. The seventh line of the example declares a variable
-\code{fraction}, which is later used with article \code{the} (the eighth
+\ff{a}. The seventh line of the example declares a variable
+\ff{fraction}, which is later used with article \ff{the} (the eighth
 line). If, by mistake, variable declaration is removed from the first step
@@
 for compliance with the class expected. The eighth line of the example
-expects method ``calculates'' of class \code{Fraction} to return an object
-of class \code{Float}. The method is still informal and is not defined in
+expects method ``calculates'' of class \ff{Fraction} to return an object
+of class \ff{Float}. The method is still informal and is not defined in
 the requirements document, but when a requirements expert decides to make it
 formal and specify its details in a new use case, she has to make sure it
-returns an instance of class \code{Float}. Otherwise, the compiler would
+returns an instance of class \ff{Float}. Otherwise, the compiler would
 complain with an error.
@@
 Every method in the object-oriented way may throw an exception, using a
-build-in method \code{Fail since}. The method has one parameter of class
-\code{Text}, which explains the reason. In the example in
-Fig.~\ref{fig:example}, method ``calculates'' of class \code{Fraction} may
+build-in method \ff{Fail since}. The method has one parameter of class
+\ff{Text}, which explains the reason. In the example in
+the Fig.~\ref{fig:example}, method ``calculates'' of class \ff{Fraction} may
 throw an exception and it is handled by an alternative flow. In the language, all
@@
 When semantic analysis is done the specification looks like an object of
-class \code{SuD} that encapsulates other objects and methods. This object
-hierarchy is convertible to UML diagrams~\cite{uml}. Specification in
-Fig.~\ref{fig:example} would be converted to the following UML diagrams
+class \ff{SuD} that encapsulates other objects and methods. This object
+hierarchy is convertible to UML diagrams~\citep{uml}. Specification in
+the Fig.~\ref{fig:example} would be converted to the following UML diagrams
 (to save space, the list is limited to the three most interesting
@@
 \begin{itemize}
-\item Class diagram, in Fig.~\ref{fig:uml-class}
-\item Object diagram, in Fig.~\ref{fig:uml-object}
-\item Sequence diagram of ``divides two numbers'' method, in Fig.~\ref{fig:uml-seq}
+\item Class diagram, in the Fig.~\ref{fig:uml-class}
+\item Object diagram, in the Fig.~\ref{fig:uml-object}
+\item Sequence diagram of ``divides two numbers'' method, in the Fig.~\ref{fig:uml-seq}
 \end{itemize}
@@
 
-Every UML diagram is a formal interoperable document in XMI~\cite{xmi} that
+Every UML diagram is a formal interoperable document in XMI~\citep{xmi} that
 can be rendered in web, in \LaTeX, in PDF, or translated to Java or any
-other object-oriented language. Fig.~\ref{fig:xmi} shows an example XMI
+other object-oriented language. The Fig.~\ref{fig:xmi} shows an example XMI
 for the class diagram.
@@
 \begin{figure}
-\inputminted[linenos,samepage,frame=lines,numbersep=4pt,fontsize=\small]{xml}{sample.xmi}
+\begin{ffcode}
+<uml:Model xmi:version="2.1"
+  xmlns:uml
+    ="http://www.eclipse.org/uml2/2.0.0/UML"
+  xmlns:xmi
+    ="http://schema.omg.org/spec/XMI/2.1">
+  <packagedElement name="SuD"
+    xmi:type="uml:Class">
+    <ownedOperation name="divides two numbers"
+      xmi:type="uml:Operation">
+      <ownedParameter direction="return"
+        xmi:type="uml:Parameter"/>
+    </ownedOperation>
+  </packagedElement>
+  <packagedElement name="User"
+    xmi:type="uml:Class">
+    <ownedOperation name="creates fraction"
+      xmi:type="uml:Operation">
+      <ownedParameter direction="return"
+        xmi:type="uml:Parameter"/>
+    </ownedOperation>
+  </packagedElement>
+  [...]
+</uml:Model>
+\end{ffcode}
 \caption{XMI representation of a UML diagram; a sample part}
@@
 
-In Fig.~\ref{fig:example}, ambiguity equals to $3/4$ since there are four
+In the Fig.~\ref{fig:example}, ambiguity equals to $3/4$ since there are four
 methods in total and three of them are defined informally.
@@
 A method is formally defined when its signature consists of one specially
-reserved verb \code{creates} and an optional list of constructor arguments,
-as in the seventh line of Fig.~\ref{fig:example}. A formally defined
+reserved verb \ff{creates} and an optional list of constructor arguments,
+as in the seventh line of the Fig.~\ref{fig:example}. A formally defined
 method doesn't have any ambiguity since it is absolutely clear what is
@@
 Total Java classes                                  & 80 & 75 & 295 \\
-\rowcolor{lightgray!50}
 Ambiguity achieved                                  & 0.80 & 0.95 & 0.68 \\
@@
 Third, developing a requirements specification with the language requires RE
-engineers to understand key principles of ``object thinking''~\cite{west04},
+engineers to understand key principles of ``object thinking''~\citep{west04},
 like ``everything is an object'' and ``objects expose behavior, not state.''

### 2021-08-01 sample.java

@@ -1,29 +1,0 @@
-class User {
-  public Fraction createsFraction() {
-  }
-  public void receivesResult(Float quotient) {
-  }
-  public void fails(String message) {
-  }
-}
-class Fraction {
-  private final float numerator;
-  private final float denumerator;
-  public Fraction(float num, float denum) {
-    this.numerator = num;
-    this.denumerator = denum;
-  }
-  public Float quotient() throws DivisionByZero {
-    }
-  }
-  class SuD {
-    private final User user;
-    public dividesTwoNumbers() {
-    Fraction fraction = this.user.createsFraction();
-    try {
-      this.user.receivesResult(fraction.quotient());
-    } catch (DivisionByZero ex) {
-      this.user.fails("denominator can't be zero");
-    }
-  }
-}

### 2021-08-01 article.tex
layout

@@
   \usetikzlibrary{calc}
-\usepackage{href-ul} % hyper links
 
 \newcommand{\nospell}[1]{#1}
+\setlength\headheight{24pt}
+\newcommand\figcaption[1]{\Description{#1}\caption{#1}}
+\raggedbottom
 
@@
 To overcome these drawbacks, a new language was introduced,
-implemented, and tested. Section~\ref{sec:Requs} of
+implemented, and tested. The Sec.~\ref{sec:Requs} of
 the article presents the language and gives a simple example of its use,
 specifying requirements of a sample software system.
-Section~\ref{sec:lexical} briefly outlines lexical rules of the language.
-Section~\ref{sec:syntax} describes the syntax using extended \nospell{Backus-Naur}
-form and gives a few practical examples. Section~\ref{sec:semantic} outlines
+The Sec.~\ref{sec:lexical} briefly outlines lexical rules of the language.
+The Sec.~\ref{sec:syntax} describes the syntax using extended \nospell{Backus-Naur}
+form and gives a few practical examples. The Sec.~\ref{sec:semantic} outlines
 a semantic layer of the compiler that is intended to detect vast majority
 of inconsistencies in a requirements document.
-Section~\ref{sec:uml} gives an example of to-UML mapping implemented
-by the compiler. Section~\ref{sec:ambiguity}
+The Sec.~\ref{sec:uml} gives an example of to-UML mapping implemented
+by the compiler.
+The Sec.~\ref{sec:ambiguity}
 presents a numeric metric that measures ambiguity in a requirements
 specification and could be used as a guidance for system analysts.
-Section~\ref{sec:results} presents empirical results obtained from three
+The Sec.~\ref{sec:results} presents empirical results obtained from three
 commercial Java web projects over the last few years.
-Section~\ref{sec:conclusion} presents conclusions and observations made by
+The Sec.~\ref{sec:conclusion} presents conclusions and observations made by
 the author.
@@
 \end{ffcode}
-\caption{Software Requirements Specification of an example math calculator}
+\figcaption{Software Requirements Specification of an example math calculator}
 \label{fig:example}
@@
 all other objects and may not be ``created'' (similar to a singleton in OOP).
-\ff{Float} is a built-in class, as well as \ff{Text}, \ff{Integer}, and
+The \ff{Float} is a built-in class, as well as the \ff{Text}, the \ff{Integer}, and
 a few others. Classes have methods, which are defined as use cases with a
@@
 
-In the example, use case \ff{UC1} is a method of class \ff{SuD} and its
+In the example, the use case \ff{UC1} is a method of the class \ff{SuD} and its
 name is ``divides two numbers.'' Every flow step in a use case can either
 \begin{inparaenum}[\itshape a\upshape)] \item call other method, or \item
-create (instantiate) an object. \end{inparaenum} In the example, step
-\ff{1} of \ff{UC1} is creating a new object \ff{fraction} of class
-\ff{Fraction}, while step \ff{2} is calling a method ``calculates'' of
-\ff{fraction} in order to get back an object named \ff{quotient} of
-class \ff{Float}.
+create (instantiate) an object. \end{inparaenum} In the example, the step
+\ff{1} of the \ff{UC1} is creating a new object \ff{fraction} of the class
+\ff{Fraction}, while the step \ff{2} is calling a method ``calculates'' of
+the \ff{fraction} in order to get back an object named \ff{quotient} of
+the class \ff{Float}.
 
@@
 class User {
-  Fraction createsFraction() {
-  }
-  void receivesResult(Float quotient) {
-  }
-  void fails(String message) {
-  }
+  Fraction createsFraction() {}
+  void receivesResult(Float quotient) {}
+  void fails(String message) {}
 }
@@
   void dividesTwoNumbers() {
-    Fraction fraction = this.user.createsFraction();
+    Fraction fraction =
+      this.user.createsFraction();
     try {
@@
 \end{ffcode}
-\caption{Java equivalent of Requs requirements document from the Fig.~\ref{fig:example}}
+\figcaption{Java equivalent of Requs requirements document from the Fig.~\ref{fig:example}}
 \label{fig:java}
@@
 
-Requs, in its current version, doesn't have any lexical/syntax constructs
+Requs, in its current version, doesn't have any lexical or syntax constructs
 for non-functional requirements (NFRs) specification. This is a subject for
@@
 
-In the Fig.~\ref{fig:example}, \ff{SuD}, \ff{Fraction}, \ff{Float}, and
+In the Fig.~\ref{fig:example}, the entities \ff{SuD}, \ff{Fraction}, \ff{Float}, and
 \ff{User} are classes. They start with capital letters and contain only
-letters (both lowercase and capital). \ff{user}, \ff{fraction}, and
-\ff{quotient} are variables bound to instances/objects of classes.
+letters (both lowercase and capital). The \ff{user}, the \ff{fraction}, and
+the \ff{quotient} are the variables bound to instances of classes.
 
@@
 certain places of the text, including \ff{the}, \ff{a}, \ff{includes},
-\ff{requires}, etc. (A full list of them is defined below in Section~\ref{sec:syntax}.)
+\ff{requires}, etc.
+A full list of them is defined below in the Sec.~\ref{sec:syntax}.
 
@@
 
-The second line in the the Fig.~\ref{fig:example} declares class \ff{Fraction}
+The second line in the Fig.~\ref{fig:example} declares the class \ff{Fraction}
 without any super-class but with an informal description
@@
 The Fig.~\ref{fig:example} defines constructor arguments only for
-\ff{Fraction} class. Class \ff{User} don't have any arguments. Class
-\ff{SuD} is a pre-defined class without arguments. Class \ff{Float} is a
+the \ff{Fraction} class. The class \ff{User} don't have any arguments. The class
+\ff{SuD} is a pre-defined class without arguments. The class \ff{Float} is a
 pre-defined class with one argument.
@@
 Class without arguments in constructor normally means that it represents an
-entity outside of system scope, like \ff{User}.
+entity outside of system scope, like the \ff{User}.
 
@@
 For traceability, all methods have unique names across the entire
-SRS starting with \ff{UC} (use case). In other words,
+SRS starting with the \ff{UC} (use case). In other words,
 methods are use cases owned by classes.
@@
 create a fraction with just one parameter or two parameters of classes other
-than \ff{Float} would be rejected by the compiler.
+than the \ff{Float} would be rejected by the compiler.
 
@@
 for compliance with the class expected. The eighth line of the example
-expects method ``calculates'' of class \ff{Fraction} to return an object
+expects method ``calculates'' of the class \ff{Fraction} to return an object
 of class \ff{Float}. The method is still informal and is not defined in
@@
 formal and specify its details in a new use case, she has to make sure it
-returns an instance of class \ff{Float}. Otherwise, the compiler would
+returns an instance of the class \ff{Float}. Otherwise, the compiler would
 complain with an error.
@@
 Every method in the object-oriented way may throw an exception, using a
-build-in method \ff{Fail since}. The method has one parameter of class
+build-in method \ff{Fail since}. The method has one parameter of the class
 \ff{Text}, which explains the reason. In the example in
-the Fig.~\ref{fig:example}, method ``calculates'' of class \ff{Fraction} may
+the Fig.~\ref{fig:example}, method ``calculates'' of the class \ff{Fraction} may
 throw an exception and it is handled by an alternative flow. In the language, all
@@
 When semantic analysis is done the specification looks like an object of
-class \ff{SuD} that encapsulates other objects and methods. This object
+the class \ff{SuD} that encapsulates other objects and methods. This object
 hierarchy is convertible to UML diagrams~\citep{uml}. Specification in
@@
 \end{tikzpicture}
-\caption{UML class diagram for the sample project}
+\figcaption{UML class diagram for the sample project}
 \label{fig:uml-class}
@@
 \end{tikzpicture}
-\caption{UML object diagram for the sample project}
+\figcaption{UML object diagram for the sample project}
 \label{fig:uml-object}
@@
 \end{sequencediagram}
-\caption{UML sequence diagram for ``divides two numbers'' method of class SuD of the sample project}
+\figcaption{UML sequence diagram for ``divides two numbers'' method of class SuD of the sample project}
 \label{fig:uml-seq}
@@
 \end{ffcode}
-\caption{XMI representation of a UML diagram; a sample part}
+\figcaption{XMI representation of a UML diagram; a sample part}
 \label{fig:xmi}
@@
 requirements until a target ambiguity is reached. In the commercial projects
-explained in Section~\ref{sec:results}, the $A$ metric was used for the planning
+explained in the Sec.~\ref{sec:results}, the $A$ metric was used for the planning
 of RE activities. Every iteration a group of RE specialists had a goal of
@@
 \end{tabular}
-\caption{Empirically collected data from three commercial projects}
+\figcaption{Empirically collected data from three commercial projects}
 \label{table:numbers}
@@
 
-\bibliographystyle{abbrv}
+\bibliographystyle{ACM-Reference-Format}
+\raggedright
 \bibliography{main}
+\clearpage
 

### 2021-09-13 article.tex
clean
polished

@@
 \title{Combining Object-Oriented Paradigm and Controlled Natural Language for Requirements Specification}
-\author{Yegor Bugayenko\\yegor256@gmail.com}
+\author{Yegor Bugayenko}{}{}
+\email{yegor256@gmail.com}
+\affiliation{\institution{}\city{Moscow}\country{Russia}}
+\ccsdesc[100]{Requirements Engineering}
+\keywords{Requirements, Natural Language Processing}
+
 \begin{abstract} Natural language is the dominant form of writing software requirements.
@@
 complete, consistent, ranked for importance and/or stability, verifiable,
-modifiable, and traceable~\citep{ieee830}. Many times, it was proven that an SRS
+modifiable, and traceable~\citep{ieee830}. Many times, it was demonstrated that an SRS
 document that lacks some of these properties has a high likelihood of
@@
 
-Natural Language (NL) is the main presentation means in industrial
+Natural Language (NL) is the main presentation mean in industrial
 requirements documents~\citep{kof10,sinha10}, mostly due to its flexibility.
@@
 Second, they all require sequential flow of language constructs, while in
-modern agile software projects wiki pages is the most popular storage for
+modern Agile software projects wiki pages is one of the most popular storage for
 requirements documentation, instead of files or databases. Wiki pages, by
@@
 To overcome these drawbacks, a new language was introduced,
-implemented, and tested. The Sec.~\ref{sec:Requs} of
+implemented, and tested.
+Sec.~\ref{sec:Requs} of
 the article presents the language and gives a simple example of its use,
 specifying requirements of a sample software system.
-The Sec.~\ref{sec:lexical} briefly outlines lexical rules of the language.
-The Sec.~\ref{sec:syntax} describes the syntax using extended \nospell{Backus-Naur}
-form and gives a few practical examples. The Sec.~\ref{sec:semantic} outlines
+Sec.~\ref{sec:lexical} briefly outlines lexical rules of the language.
+Sec.~\ref{sec:syntax} describes the syntax using extended \nospell{Backus-Naur}
+form and gives a few practical examples.
+Sec.~\ref{sec:semantic} outlines
 a semantic layer of the compiler that is intended to detect vast majority
 of inconsistencies in a requirements document.
-The Sec.~\ref{sec:uml} gives an example of to-UML mapping implemented
+Sec.~\ref{sec:uml} gives an example of to-UML mapping implemented
 by the compiler.
-The Sec.~\ref{sec:ambiguity}
+Sec.~\ref{sec:ambiguity}
 presents a numeric metric that measures ambiguity in a requirements
 specification and could be used as a guidance for system analysts.
-The Sec.~\ref{sec:results} presents empirical results obtained from three
+Sec.~\ref{sec:results} presents empirical results obtained from three
 commercial Java web projects over the last few years.
-The Sec.~\ref{sec:conclusion} presents conclusions and observations made by
+Sec.~\ref{sec:conclusion} presents conclusions and observations made by
 the author.
@@
 can be read and understood by compilers, system analysts, end-users, and
-business people. The Fig.~\ref{fig:example} presents an example of a short
+business people. Fig.~\ref{fig:example} presents an example of a short
 requirements definition document that describes a calculator that asks its
@@
 
-The Fig.~\ref{fig:java} contains a close equivalent program in Java.
+Fig.~\ref{fig:java} contains a close equivalent program in Java.
 
@@
 
-\begin{grammar}
-<class> ::= `[A-Z][A-Za-z]+'
+\begin{ffcode*}{linenos=false}
+<class> ::= /[A-Z][A-Za-z]+/
+<word> ::= /[a-z]+/
+<use-case-ID> ::= /UC[0-9\\.]+/
+<flow-ID> ::= /[0-9]+/
+<informal> ::= /"([^"]|$\vert$|\\")+"/
+\end{ffcode*}
 
-<word> ::= `[a-z]+'
-
-<use-case-ID> ::= `UC[0-9\\.]+'
-
-<flow-ID> ::= `[0-9]+'
-
-<informal> ::= `"([^"]|\\")+"'
-\end{grammar}
-
-In the Fig.~\ref{fig:example}, the entities \ff{SuD}, \ff{Fraction}, \ff{Float}, and
+In Fig.~\ref{fig:example}, the entities \ff{SuD}, \ff{Fraction}, \ff{Float}, and
 \ff{User} are classes. They start with capital letters and contain only
@@
 \ff{requires}, etc.
-A full list of them is defined below in the Sec.~\ref{sec:syntax}.
+A full list of them is defined below in Sec.~\ref{sec:syntax}.
 
@@
 
-\begin{grammar}
+\begin{ffcode*}{linenos=false}
 <SRS> ::= <clause>+
-
-<clause> ::= <class-declaration> \\
-| <class-construction> \\
-| <method-declaration> \\
-| <alternative-flow-declaration>
-\end{grammar}
+<clause> ::= <class-declaration>
+  |$\vert$| <class-construction>
+  |$\vert$| <method-declaration>
+  |$\vert$| <alternative-flow-declaration>
+\end{ffcode*}
 
@@
 
-\begin{grammar}
-<class-declaration> ::= <class> <is-a> <parent-class>
+\begin{ffcode*}{linenos=false}
+<class-declaration> ::= <class>
+  <is-a> <parent-class>
+<parent-class> ::= <informal> |$\vert$| <class>
+<is-a> ::= `is' ( `a' |$\vert$| `an' )
+\end{ffcode*}
 
-<parent-class> ::= <informal> | <class>
-
-<is-a> ::= `is' ( `a' | `an' )
-\end{grammar}
-
-The second line in the Fig.~\ref{fig:example} declares the class \ff{Fraction}
+The second line in Fig.~\ref{fig:example} declares the class \ff{Fraction}
 without any super-class but with an informal description
@@
 
-\begin{grammar}
-<class-construction> ::= <class> <includes> `:' <slots>
+\begin{ffcode*}{linenos=false}
+<class-construction> ::= <class> <includes>
+  `:' <slots>
+<includes> ::= `includes' |$\vert$| `needs'
+<slots> ::= <slot> ( `,' <slot> )*
+<slot> ::= <variable> ( `as' <class> )?
+<variable> ::= <word> `-s'?
+\end{ffcode*}
 
-<includes> ::= `includes' | `needs'
-
-<slots> ::= <slot> ( `,' <slot> )*
-
-<slot> ::= <variable> ( `as' <class> )?
-
-<variable> ::= <word> `-s'?
-\end{grammar}
-
-The Fig.~\ref{fig:example} defines constructor arguments only for
+Fig.~\ref{fig:example} defines constructor arguments only for
 the \ff{Fraction} class. The class \ff{User} don't have any arguments. The class
@@
 
-\begin{grammar}
-<method-declaration> ::= <use-case-ID> `where' \\
+\begin{ffcode*}{linenos=false}
+<method-declaration> ::= <use-case-ID> `where'
 <class> <signature> `:' <flows>
-
-<signature> ::= ( <word>+ | <informal> ) \\
+<signature> ::= ( <word>+ |$\vert$| <informal> )
   <subject>? <using>?
-
-<subject> ::= <class> <binding>? | `the' <variable>
-
+<subject> ::= <class> <binding>?
+  |$\vert$| `the' <variable>
 <binding> ::= `(' `a' <variable> `)'
-
-<using> ::= `using' <subject> ( `and' <subject> )*
-
+<using> ::= `using' <subject>
+  ( `and' <subject> )*
 <flows> ::= <flow> ( `;' <flow> )+
-
-<flow> ::= <flow-ID> `.' \\
-( `The' <variable> <signature>
-| `Fail' `since' <informal> )
-\end{grammar}
+<flow> ::= <flow-ID> `.'
+  (
+    `The' <variable> <signature>
+    |$\vert$|
+    `Fail' `since' <informal>
+  )
+\end{ffcode*}
 
@@
 
-\begin{grammar}
-<alternative-flow-declaration> ::= <use-case-ID> `/' \\
-<flow-ID> `when' <informal> `:' <flows>
-\end{grammar}
+\begin{ffcode*}{linenos=false}
+<alternative-flow-declaration> ::= <use-case-ID> `/' <flow-ID> `when' <informal> `:' <flows>
+\end{ffcode*}
 
@@
 \begin{inparaenum}[\itshape a\upshape)]
-\item be informal (the 8th, 9th, and 11th lines in the example at the Fig.~\ref{fig:example}),
+\item be informal (the 8th, 9th, and 11th lines in the example at Fig.~\ref{fig:example}),
 \item formally refer to the method declared in a class, or
@@
 \ff{Text}, which explains the reason. In the example in
-the Fig.~\ref{fig:example}, method ``calculates'' of the class \ff{Fraction} may
+Fig.~\ref{fig:example}, method ``calculates'' of the class \ff{Fraction} may
 throw an exception and it is handled by an alternative flow. In the language, all
@@
 hierarchy is convertible to UML diagrams~\citep{uml}. Specification in
-the Fig.~\ref{fig:example} would be converted to the following UML diagrams
+Fig.~\ref{fig:example} would be converted to the following UML diagrams
 (to save space, the list is limited to the three most interesting
@@
 \begin{itemize}
-\item Class diagram, in the Fig.~\ref{fig:uml-class}
-\item Object diagram, in the Fig.~\ref{fig:uml-object}
-\item Sequence diagram of ``divides two numbers'' method, in the Fig.~\ref{fig:uml-seq}
+\item Class diagram, in Fig.~\ref{fig:uml-class}
+\item Object diagram, in Fig.~\ref{fig:uml-object}
+\item Sequence diagram of ``divides two numbers'' method, in Fig.~\ref{fig:uml-seq}
 \end{itemize}
@@
 can be rendered in web, in \LaTeX, in PDF, or translated to Java or any
-other object-oriented language. The Fig.~\ref{fig:xmi} shows an example XMI
+other object-oriented language. Fig.~\ref{fig:xmi} shows an example XMI
 for the class diagram.
@@
 
-In the Fig.~\ref{fig:example}, ambiguity equals to $3/4$ since there are four
+In Fig.~\ref{fig:example}, ambiguity equals to $3/4$ since there are four
 methods in total and three of them are defined informally.
@@
 reserved verb \ff{creates} and an optional list of constructor arguments,
-as in the seventh line of the Fig.~\ref{fig:example}. A formally defined
+as in the seventh line of Fig.~\ref{fig:example}. A formally defined
 method doesn't have any ambiguity since it is absolutely clear what is
@@
 requirements until a target ambiguity is reached. In the commercial projects
-explained in the Sec.~\ref{sec:results}, the $A$ metric was used for the planning
+explained in Sec.~\ref{sec:results}, the $A$ metric was used for the planning
 of RE activities. Every iteration a group of RE specialists had a goal of

### 2021-09-13 article.tex
formatting

@@
 \raggedbottom
+\tolerance=2000
 
@@
 
-Fig.~\ref{fig:example} defines constructor arguments only for
+Fig.~\ref{fig:example} defines the arguments of the constructor only for
 the \ff{Fraction} class. The class \ff{User} don't have any arguments. The class
@@
 
-Third, developing a requirements specification with the language requires RE
+Third, developing requirements specification with the language requires RE
 engineers to understand key principles of ``object thinking''~\citep{west04},

### 2021-09-17 article.tex
bcnc

@@
-\documentclass[sigplan,10pt,nonacm=true]{acmart}
-\settopmatter{printfolios=false,printccs=false,printacmref=false}
+\documentclass[sigplan,10pt]{acmart}
+% \settopmatter{printfolios=false,printccs=false,printacmref=false}
 \usepackage[underline=true,rounded corners=false]{pgf-umlsd} % for UML sequence diagrams
@@
 \affiliation{\institution{}\city{Moscow}\country{Russia}}
-\ccsdesc[100]{Requirements Engineering}
+\setcopyright{acmcopyright}
+\copyrightyear{2021}
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011074.10011075.10011076</concept_id>
+<concept_desc>Software and its engineering~Requirements analysis</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+\ccsdesc[500]{Software and its engineering~Requirements analysis}
+\acmConference[BCNC 2021]
+  {1st ACM SIGPLAN International Workshop on Beyond Code: No Code}
+  {October 17, 2021}
+  {Chicago, IL, US}
 \keywords{Requirements, Natural Language Processing}
@@
 \begin{ffcode*}{linenos=false}
-<alternative-flow-declaration> ::= <use-case-ID> `/' <flow-ID> `when' <informal> `:' <flows>
+<alternative-flow-declaration> ::=
+  <use-case-ID> `/' <flow-ID>
+  `when' <informal> `:' <flows>
 \end{ffcode*}

### 2021-09-17 article.tex
doi
fin

@@
-\documentclass[sigplan,10pt]{acmart}
+\documentclass[sigplan,10pt,screen]{acmart}
 % \settopmatter{printfolios=false,printccs=false,printacmref=false}
@@
 \usepackage{ffcode}
+\usepackage{flushend}
 \usepackage{tikz}
@@
 \title{Combining Object-Oriented Paradigm and Controlled Natural Language for Requirements Specification}
-\author{Yegor Bugayenko}{}{}
+\author{Yegor Bugayenko}
+\orcid{0000-0001-6370-0678}
 \email{yegor256@gmail.com}
-\affiliation{\institution{}\city{Moscow}\country{Russia}}
+\affiliation{\city{Moscow}\country{Russia}}
+
 \setcopyright{acmcopyright}
+\acmPrice{15.00}
+\acmDOI{10.1145/3486949.3486963}
+\acmYear{2021}
 \copyrightyear{2021}
+\acmSubmissionID{splashws21bcncmain-p41-p}
+\acmISBN{978-1-4503-9125-2/21/10}
+\acmConference[BCNC '21]{Proceedings of the 1st ACM SIGPLAN International Workshop on Beyond Code: No Code}{October 17, 2021}{Chicago, IL, USA}
+\acmBooktitle{Proceedings of the 1st ACM SIGPLAN International Workshop on Beyond Code: No Code (BCNC '21), October 17, 2021, Chicago, IL, USA}
+
 \begin{CCSXML}
@@
 \ccsdesc[500]{Software and its engineering~Requirements analysis}
-\acmConference[BCNC 2021]
-  {1st ACM SIGPLAN International Workshop on Beyond Code: No Code}
-  {October 17, 2021}
-  {Chicago, IL, US}
 \keywords{Requirements, Natural Language Processing}
@@
 Third, developing requirements specification with the language requires RE
-engineers to understand key principles of ``object thinking''~\citep{west04},
+engineers to understand key principles of ``object thinking''~\citep{west04,eo1},
 like ``everything is an object'' and ``objects expose behavior, not state.''

### 2021-12-18 article.tex
links

@@
-\documentclass[sigplan,10pt,screen]{acmart}
-% \settopmatter{printfolios=false,printccs=false,printacmref=false}
+\documentclass[sigplan,screen]{acmart}
 \usepackage[underline=true,rounded corners=false]{pgf-umlsd} % for UML sequence diagrams
-\usepackage[square]{syntax} % for BNF grammar
-  \setlength{\grammarparsep}{1pt}
-  \setlength{\grammarindent}{2em}
 \usepackage{array} % complex structure of TABULAR-s
 \usepackage{paralist} % for inline itemized lists env{inparaenum}
-\usepackage{natbib} % for customization of bibliography headers
-\usepackage{ffcode}
-\usepackage{flushend}
+\usepackage[nonumbers,noframes]{ffcode}
 \usepackage{tikz}
@@
 \newcommand{\nospell}[1]{#1}
-\setlength\headheight{24pt}
 \newcommand\figcaption[1]{\Description{#1}\caption{#1}}
-\raggedbottom
-\tolerance=2000
+% \tolerance=2000
 
@@
 \email{yegor256@gmail.com}
-\affiliation{\city{Moscow}\country{Russia}}
+\affiliation{\institution{Huawei}\city{Moscow}\country{Russia}}
 
@@
 
-It is an object-oriented CNL that resembles English and, because of that,
+Requs\footnote{\url{https://www.requs.org}}
+is an object-oriented CNL that resembles English and, because of that,
 can be read and understood by compilers, system analysts, end-users, and
@@
 
-\subsection{Method calls match declarations}
+\subsection{Method Calls Match Declarations}
 
@@
 
-\subsection{Object creating have necessary constructor arguments}
+\subsection{Object Creating Have Necessary Constructor Arguments}
 
@@
 
-\subsection{Variable bindings precede their usage}
+\subsection{Variable Bindings Precede Their Usage}
 
@@
 
-\subsection{Methods return objects of required classes}
+\subsection{Methods Return Objects of Required Classes}
 
@@
 
-\subsection{Failures are handled by alternative flows}
+\subsection{Failures Are Handled by Alternative Flows}
 
@@
 \begin{table}
+\figcaption{Empirically collected data from three commercial projects}
+\label{table:numbers}
 \begin{tabular}{>{\raggedright}p{14em}rrr}
@@
 \end{tabular}
-\figcaption{Empirically collected data from three commercial projects}
-\label{table:numbers}
 \end{table}
@@
 
+%%% -*-BibTeX-*-
+%%% Do NOT edit. File created by BibTeX with style
+%%% ACM-Reference-Format-Journals [18-Jan-2012].
+
+\begin{thebibliography}{18}
+
+%%% ====================================================================
+%%% NOTE TO THE USER: you can override these defaults by providing
+%%% customized versions of any of these macros before the \bibliography
+%%% command.  Each of them MUST provide its own final punctuation,
+%%% except for \shownote{}, \showDOI{}, and \showURL{}.  The latter two
+%%% do not use final punctuation, in order to avoid confusing it with
+%%% the Web address.
+%%%
+%%% To suppress output of a particular field, define its macro to expand
+%%% to an empty string, or better, \unskip, like this:
+%%%
+%%% \newcommand{\showDOI}[1]{\unskip}   % LaTeX syntax
+%%%
+%%% \def \showDOI #1{\unskip}           % plain TeX syntax
+%%%
+%%% ====================================================================
+
+\ifx \showCODEN    \undefined \def \showCODEN     #1{\unskip}     \fi
+\ifx \showDOI      \undefined \def \showDOI       #1{#1}\fi
+\ifx \showISBNx    \undefined \def \showISBNx     #1{\unskip}     \fi
+\ifx \showISBNxiii \undefined \def \showISBNxiii  #1{\unskip}     \fi
+\ifx \showISSN     \undefined \def \showISSN      #1{\unskip}     \fi
+\ifx \showLCCN     \undefined \def \showLCCN      #1{\unskip}     \fi
+\ifx \shownote     \undefined \def \shownote      #1{#1}          \fi
+\ifx \showarticletitle \undefined \def \showarticletitle #1{#1}   \fi
+\ifx \showURL      \undefined \def \showURL       {\relax}        \fi
+% The following commands are used for tagged output and should be
+% invisible to TeX
+\providecommand\bibfield[2]{#2}
+\providecommand\bibinfo[2]{#2}
+\providecommand\natexlab[1]{#1}
+\providecommand\showeprint[2][]{arXiv:#2}
+
+\bibitem[\protect\citeauthoryear{??}{ste}{2005}]%
+        {ste05}
+ \bibinfo{year}{2005}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{ASD Simplified Technical English.
+  International specification for the preparation of maintenance documentation
+  in a controlled language}}.
+\newblock Number Specification ASD-STE100. \bibinfo{publisher}{Simplified
+  Technical English Maintenance Group (STEMG)}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{??}{uml}{2005}]%
+        {uml}
+ \bibinfo{year}{2005}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{Unified Modeling Language:
+  Superstructure}}.
+\newblock \bibinfo{type}{{T}echnical {R}eport}. \bibinfo{institution}{Object
+  Management Group}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{??}{pmb}{2008}]%
+        {pmbok4}
+ \bibinfo{year}{2008}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{Project Management Body of Knowledge,
+  Guide} (\bibinfo{edition}{4th} ed.)}.
+\newblock \bibinfo{publisher}{Project Management Institute}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Bugayenko}{Bugayenko}{2016}]%
+        {eo1}
+\bibfield{author}{\bibinfo{person}{Yegor Bugayenko}.}
+  \bibinfo{year}{2016}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{Elegant Objects}}.
+  Vol.~\bibinfo{volume}{1}.
+\newblock \bibinfo{publisher}{Amazon}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Chen}{Chen}{2009}]%
+        {chen09}
+\bibfield{author}{\bibinfo{person}{Betty~H.C. Chen}.}
+  \bibinfo{year}{2009}\natexlab{}.
+\newblock \showarticletitle{Current and Future Research Directions in
+  Requirements Engineering}. In \bibinfo{booktitle}{\emph{Design Requirements
+  Engineering: A Ten-Year Perspective In Design Requirements Engineering: A
+  Ten-Year Perspective}}, Vol.~\bibinfo{volume}{14}. \bibinfo{pages}{11--43}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Clark, Harrison, Jenkins, Thompson, and
+  Wo-jcik}{Clark et~al\mbox{.}}{2005}]%
+        {clark05}
+\bibfield{author}{\bibinfo{person}{P. Clark}, \bibinfo{person}{P. Harrison},
+  \bibinfo{person}{T. Jenkins}, \bibinfo{person}{J. Thompson}, {and}
+  \bibinfo{person}{R.~H. Wo-jcik}.} \bibinfo{year}{2005}\natexlab{}.
+\newblock \showarticletitle{Acquiring and Using World Knowledge Using a
+  Restricted Subset of English}. In \bibinfo{booktitle}{\emph{FLAIRS}}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Cockburn}{Cockburn}{2001}]%
+        {cockburn01}
+\bibfield{author}{\bibinfo{person}{Alistair Cockburn}.}
+  \bibinfo{year}{2001}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{Writing Effective Use Cases}}.
+\newblock \bibinfo{publisher}{Addison-Wesley}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Fuchs, Kaljurand, and Schneider}{Fuchs
+  et~al\mbox{.}}{2006}]%
+        {ace06}
+\bibfield{author}{\bibinfo{person}{Norbert~E. Fuchs}, \bibinfo{person}{Kaarel
+  Kaljurand}, {and} \bibinfo{person}{Gerold Schneider}.}
+  \bibinfo{year}{2006}\natexlab{}.
+\newblock \showarticletitle{{Attempto Controlled English Meets the Challenges
+  of Knowledge Representation, Reasoning, Interoperability and User
+  Interfaces}}. In \bibinfo{booktitle}{\emph{FLAIRS}}. Department of
+  Informatics \& Institute of Computational Linguistics.
+\newblock
+
+\bibitem[\protect\citeauthoryear{IEEE}{IEEE}{1998}]%
+        {ieee830}
+\bibfield{author}{\bibinfo{person}{IEEE}.} \bibinfo{year}{1998}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{Recommended Practice for Software
+  Requirements Specifications}}.
+\newblock \bibinfo{type}{{T}echnical {R}eport} IEEE Std 830-1998.
+  \bibinfo{institution}{The Institute of Electrical and Electronics Engineers}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Kof}{Kof}{2010}]%
+        {kof10}
+\bibfield{author}{\bibinfo{person}{L. Kof}.} \bibinfo{year}{2010}\natexlab{}.
+\newblock \showarticletitle{From Requirements Documents to System Models: A
+  Tool for Interactive Semi-Automatic Translation}. In
+  \bibinfo{booktitle}{\emph{18th IEEE International Requirements Engineering
+  Conference (RE)}}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Mavin, Wilkinson, Harwood, and Novak}{Mavin
+  et~al\mbox{.}}{2009}]%
+        {ears09}
+\bibfield{author}{\bibinfo{person}{Alistair Mavin}, \bibinfo{person}{Philip
+  Wilkinson}, \bibinfo{person}{Adrian Harwood}, {and} \bibinfo{person}{Mark
+  Novak}.} \bibinfo{year}{2009}\natexlab{}.
+\newblock \showarticletitle{{EARS (Easy Approach to Requirements Syntax)}}. In
+  \bibinfo{booktitle}{\emph{17th IEEE International Requirements Engineering
+  Conference ({RE})}}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Moynihan}{Moynihan}{2000}]%
+        {moynihan00}
+\bibfield{author}{\bibinfo{person}{Tony Moynihan}.}
+  \bibinfo{year}{2000}\natexlab{}.
+\newblock \showarticletitle{`Requirements-uncertainty': should it be a latent,
+  aggregate or profile construct?}. In \bibinfo{booktitle}{\emph{Australian
+  Software Engineering Conference}}. \bibinfo{pages}{181--188}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{OMG}{OMG}{2011}]%
+        {xmi}
+\bibfield{author}{\bibinfo{person}{OMG}.} \bibinfo{year}{2011}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{{MOF} 2.0/{XMI} Mapping, Version 2.4.1}}.
+\newblock \bibinfo{type}{{T}echnical {R}eport}. \bibinfo{institution}{Object
+  Management Group}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{Schwitter}{Schwitter}{[n.\,d.]}]%
+        {peng}
+\bibfield{author}{\bibinfo{person}{Rolf Schwitter}.}
+  \bibinfo{year}{[n.\,d.]}\natexlab{}.
+\newblock \bibinfo{title}{Processable {ENGlish}}.
+\newblock
+\newblock
+\urldef\tempurl%
+\url{http://web.science.mq.edu.au/~rolfs/peng/}
+\showURL{%
+\tempurl}
+
+\bibitem[\protect\citeauthoryear{Sinha, Paradkar, Takeuchi, and Nakamura}{Sinha
+  et~al\mbox{.}}{2010}]%
+        {sinha10}
+\bibfield{author}{\bibinfo{person}{A. Sinha}, \bibinfo{person}{A. Paradkar},
+  \bibinfo{person}{H. Takeuchi}, {and} \bibinfo{person}{T. Nakamura}.}
+  \bibinfo{year}{2010}\natexlab{}.
+\newblock \showarticletitle{Extending Automated Analysis of Natural Language
+  Use Cases to Other Languages}. In \bibinfo{booktitle}{\emph{18th IEEE
+  International Requirements Engineering Conference (RE), 2010}}.
+\newblock
+
+\bibitem[\protect\citeauthoryear{West}{West}{2004}]%
+        {west04}
+\bibfield{author}{\bibinfo{person}{David West}.}
+  \bibinfo{year}{2004}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{Object Thinking}}.
+\newblock \bibinfo{publisher}{Microsoft Press}.
+\newblock
+
+\newpage
+
+\bibitem[\protect\citeauthoryear{Wiegers}{Wiegers}{2003}]%
+        {wiegers03}
+\bibfield{author}{\bibinfo{person}{Karl Wiegers}.}
+  \bibinfo{year}{2003}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{Software Requirements, Second Edition}}.
+\newblock \bibinfo{publisher}{{Microsoft Press}}.
+\newblock
+\showISBNx{0735618798}
+
+\bibitem[\protect\citeauthoryear{Woodcock and Davies}{Woodcock and
+  Davies}{1996}]%
+        {woodcock96}
+\bibfield{author}{\bibinfo{person}{J. Woodcock} {and} \bibinfo{person}{J.
+  Davies}.} \bibinfo{year}{1996}\natexlab{}.
+\newblock \bibinfo{booktitle}{\emph{Using Z-Specification, Refinement and
+  Proof}}.
+\newblock \bibinfo{publisher}{Prentice Hall}.
+\newblock
+
+\end{thebibliography}
 \bibliographystyle{ACM-Reference-Format}
-\raggedright
-\bibliography{main}
-\clearpage
 
